### 2025-11-12 time-series-analysis.qmd
20251111 - components of time series

>>>>

There are four (potential) key components of time series data:

- seasonality
- trend
- cycle
- irregularity

*Seasonality* corresponds to fluctuations that occur at fixed and known periods (e.g., time of day, day of week, or month of year) [@Hyndman2021].
A *trend* represents a long-term change (i.e., increase or decrease) in the series.
A *cycle* reflects one or more rises and falls that are not of a fixed frequency, with durations typically lasting at least two years [@Hyndman2021].
*Irregularity* is random variation or noise.
